PANCHANGE_data_preprocessing

### 2020-08-31 PANCHANGE_data_preprocessing2020-07-15.Rmd
added prefer not to answer and dont know to mhd control

ramp.mhd.raw.id <- ramp.mhd.raw %>%
  drop_na(`Login ID`) %>% # Drop NAs
  distinct(`Login ID`, .keep_all = TRUE) %>% # Remove duplicates based on ID
  separate(`Login ID`, into = c("Sample", "ID"), sep = 4) %>% # Split ID in Sample and Number
  separate(ID, into = "ID", sep = 7) %>%
  mutate(
    ID = as.numeric(ID)) %>%
  select(
    Sample, # Sample
    ID, # ID
    mhd.depression = NA.30.depression,
    mhd.pregnancy_depression = NA.30.pregnancy_depression,
    mhd.premenstrual_dysphoric_disorder_pmdd = NA.30.premenstrual_dysphoric_disorder_pmdd,
    mhd.mania_hypomania_bipolar_or_manicdepression = NA.30.mania_hypomania_bipolar_or_manicdepression,
    mhd.anxiety_nerves_or_generalised_anxiety_disorder = NA.30.anxiety_nerves_or_generalised_anxiety_disorder,
    mhd.social_anxiety_or_social_phobia = NA.30.social_anxiety_or_social_phobia,
    mhd.specific_phobia_e.g._phobia_of_flying = NA.30.specific_phobia_e.g._phobia_of_flying,
    mhd.agoraphobia = NA.30.agoraphobia,
    mhd.panic_disorder = NA.30.panic_disorder,
    mhd.panic_attacks = NA.30.panic_attacks,
    mhd.posttraumatic_stress_disorder_ptsd = NA.30.posttraumatic_stress_disorder_ptsd,
    mhd.obsessivecompulsive_disorder_ocd = NA.30.obsessivecompulsive_disorder_ocd, 
    mhd.body_dysmorphic_disorder_bdd = NA.30.body_dysmorphic_disorder_bdd,
    mhd.skin_picking_obsessive_compulsive = NA.30.skin_picking_obsessive_compulsive,
    mhd.anorexia_nervosa = NA.30.anorexia_nervosa,
    mhd.atypical_anorexia_nervosa = NA.30.atypical_anorexia_nervosa,
    mhd.bulimia_nervosa = NA.30.bulimia_nervosa,
    mhd.bingeeating_disorder = NA.30.psychological_overeating_or_bingeeating_disorder,
    mhd.schizophrenia = NA.30.schizophrenia,
    mhd.schizoaffective_disorder = NA.30.schizoaffective_disorder,
    mhd.type_psychosis_psychotic_illness = NA.30.psychosis_type_psychotic_illness,
    mhd.personality_disorder = NA.30.personality_disorder,
    mhd.autism_aspergers_or_autistic_spectrum_disorder = NA.30.autism_aspergers_or_autistic_spectrum_disorder,
    mhd.attention_deficit_hyperactivity_disorder = NA.30.attention_deficit_hyperactivity_disorder,
    mhd.personality_disorder_diagnosed = NA.30.personality_disorder_diagnosed,
    mhd.none_of_the_above = NA.30.none_of_these,
    mhd.none_of_the_above.1 = NA.30.none_of_these.1
  ) %>%
  add_numeric(., exclude = exclude_cols) %>%
  mutate_if(is.numeric, ~na_if(., -88)) %>% # Recode missing values to NAs in numeric variables
  mutate_if(is.numeric, ~na_if(., -99)) %>%
  mutate_if(is.numeric, ~na_if(., -77)) %>%
  mutate_if(is.factor, ~na_if(., "Seen but not answered")) %>% # Recode missing values to NAs in factor variables
  mutate_if(is.factor, ~na_if(., "Don't know")) %>%
  mutate_if(is.factor, ~na_if(., "Prefer not to say")) %>%
  mutate_if(is.factor, ~na_if(., "Prefer not to answer")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Seen but not answered")) %>% # Drop empty factor levels
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Don't know")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to answer")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to say"))

@@

Agoraphonia and panic disorder group
```{r mhd agoraphobia_panic_disorder}
if(GLAD == TRUE | EDGI == TRUE | NBR == TRUE | RAMP == TRUE) {
mhd.raw.id <- mhd.raw.id %>%
  mutate(
    agoraphobia_panic_disorder_numeric =
      case_when(
        mhd.agoraphobia_numeric == "1" ~ 1,
        mhd.panic_disorder_numeric == "1" ~ 1,
        mhd.panic_attacks_numeric == "1" ~ 1,
        (mhd.agoraphobia_numeric == "0" &
        mhd.panic_disorder_numeric == "0" &
        mhd.panic_attacks_numeric == "0") ~ 0
      )
  )

@@

### Controls
These individuals have no self report diagnosis of any kind that we measure. 
```{r mhd control}
mhd.raw.id <- mhd.raw.id %>%
  mutate(
    control_numeric =
      if_else(
    depressive_disorders_numeric == "0" &
    anxiety_disorders_numeric == "0" &
    eating_disorders_numeric == "0" &
    obsessive_compulsive_disorders_numeric == "0" &
    psychotic_disorders_numeric == "0" &
    mhd.mania_hypomania_bipolar_or_manicdepression_numeric == "0" &
    mhd.posttraumatic_stress_disorder_ptsd_numeric == "0" &
    autism_spectrum_disorder_numeric == "0" &
    mhd.attention_deficit_hyperactivity_disorder_numeric == "0" &
    mhd.personality_disorder_numeric == "0", 
    true = 1, 
    false = 0, 
    NA_real_)
  )

@@

```{r}
mhd.raw.id %>%
  select(
    control,
    mhd.none_of_the_above,
    mhd.none_of_the_above.1
  )

mhd.raw.id %>%
  group_by(mhd.none_of_the_above.1) %>%
  count(
    control,
    mhd.none_of_the_above
  )

mhd.raw.id %>%
  count(
    control,
    age_
  )
```

```{r Check personality_cluster_c}
if(GLAD == TRUE | NBR == TRUE | RAMP == TRUE) {
mhd.raw.id %>%
  select(
    control,
    depressive_disorders_numeric,
    anxiety_disorders_numeric,
    eating_disorders_numeric,
    obsessive_compulsive_disorders_numeric,
    psychotic_disorders_numeric,
    mhd.mania_hypomania_bipolar_or_manicdepression_numeric,
    mhd.posttraumatic_stress_disorder_ptsd_numeric,
    autism_spectrum_disorder_numeric,
    mhd.attention_deficit_hyperactivity_disorder_numeric,
    mhd.personality_disorder_numeric, 
  )
}
```

@@

Create new dataframe with the derived categories
```{r mhd categories}
if(GLAD == TRUE | NBR == TRUE | RAMP == TRUE) {
mhd.raw.categories <- mhd.raw.id %>%
  select(
    Sample,
    ID,
    anxiety_disorders_numeric,
    anxiety_disorders,
    mhd.specific_phobia_e.g._phobia_of_flying_numeric,
    mhd.specific_phobia_e.g._phobia_of_flying,
    mhd.social_anxiety_or_social_phobia_numeric,
    mhd.social_anxiety_or_social_phobia,
    agoraphobia_panic_disorder_numeric,
    agoraphobia_panic_disorder,
    depressive_disorders_numeric,
    depressive_disorders,
    eating_disorders_numeric,
    eating_disorders,
    obsessive_compulsive_disorders_numeric,
    obsessive_compulsive_disorders,
    psychotic_disorders_numeric,
    psychotic_disorders,
    mhd.mania_hypomania_bipolar_or_manicdepression_numeric,
    mhd.mania_hypomania_bipolar_or_manicdepression,
    mhd.posttraumatic_stress_disorder_ptsd_numeric,
    mhd.posttraumatic_stress_disorder_ptsd,
    autism_spectrum_disorder_numeric,
    autism_spectrum_disorder,
    mhd.attention_deficit_hyperactivity_disorder_numeric,
    mhd.attention_deficit_hyperactivity_disorder,
    mhd.personality_disorder_numeric,
    mhd.personality_disorder,
    personality_cluster_a_numeric,
    personality_cluster_a,
    personality_cluster_b_numeric,
    personality_cluster_b,
    personality_cluster_c_numeric,
    personality_cluster_c,
    depression_and_anxiety_numeric,
    depression_and_anxiety,
    restricting_vs_binge_eating_eating_disorders_numeric,
    restricting_vs_binge_eating_eating_disorders,
    bipolar_and_schizophrenia_numeric,
    bipolar_and_schizophrenia,
    control_numeric,
    control,
    disorders_total_count,
    comorbidity_total_count_numeric,
    comorbidity_total_count_factor_numeric,
    comorbidity_total_count_factor
    )
}
```

@@

```{r rename dataset for ramp}
if(RAMP == TRUE){
#there is no sex variable in ramp
dat.no.sex <- dat.raw
}
```

@@

>>>>

ramp.mhd.raw.id <- ramp.mhd.raw %>%
  drop_na(`Login ID`) %>% # Drop NAs
  distinct(`Login ID`, .keep_all = TRUE) %>% # Remove duplicates based on ID
  separate(`Login ID`, into = c("Sample", "ID"), sep = 4) %>% # Split ID in Sample and Number
  separate(ID, into = "ID", sep = 7) %>%
  mutate(
    ID = as.numeric(ID)) %>%
  select(
    Sample, # Sample
    ID, # ID
    mhd.depression = NA.30.depression,
    mhd.pregnancy_depression = NA.30.pregnancy_depression,
    mhd.premenstrual_dysphoric_disorder_pmdd = NA.30.premenstrual_dysphoric_disorder_pmdd,
    mhd.mania_hypomania_bipolar_or_manicdepression = NA.30.mania_hypomania_bipolar_or_manicdepression,
    mhd.anxiety_nerves_or_generalised_anxiety_disorder = NA.30.anxiety_nerves_or_generalised_anxiety_disorder,
    mhd.social_anxiety_or_social_phobia = NA.30.social_anxiety_or_social_phobia,
    mhd.specific_phobia_e.g._phobia_of_flying = NA.30.specific_phobia_e.g._phobia_of_flying,
    mhd.agoraphobia = NA.30.agoraphobia,
    mhd.panic_disorder = NA.30.panic_disorder,
    mhd.panic_attacks = NA.30.panic_attacks,
    mhd.posttraumatic_stress_disorder_ptsd = NA.30.posttraumatic_stress_disorder_ptsd,
    mhd.obsessivecompulsive_disorder_ocd = NA.30.obsessivecompulsive_disorder_ocd, 
    mhd.body_dysmorphic_disorder_bdd = NA.30.body_dysmorphic_disorder_bdd,
    mhd.skin_picking_obsessive_compulsive = NA.30.skin_picking_obsessive_compulsive,
    mhd.anorexia_nervosa = NA.30.anorexia_nervosa,
    mhd.atypical_anorexia_nervosa = NA.30.atypical_anorexia_nervosa,
    mhd.bulimia_nervosa = NA.30.bulimia_nervosa,
    mhd.bingeeating_disorder = NA.30.psychological_overeating_or_bingeeating_disorder,
    mhd.schizophrenia = NA.30.schizophrenia,
    mhd.schizoaffective_disorder = NA.30.schizoaffective_disorder,
    mhd.type_psychosis_psychotic_illness = NA.30.psychosis_type_psychotic_illness,
    mhd.personality_disorder = NA.30.personality_disorder,
    mhd.autism_aspergers_or_autistic_spectrum_disorder = NA.30.autism_aspergers_or_autistic_spectrum_disorder,
    mhd.attention_deficit_hyperactivity_disorder = NA.30.attention_deficit_hyperactivity_disorder,
    mhd.personality_disorder_diagnosed = NA.30.personality_disorder_diagnosed,
    mhd.none_of_the_above = NA.30.none_of_these,
    mhd.dont_know = NA.30.dont_know,
    mhd.prefer_not_to_answer = NA.30.prefer_not_to_answer,
    mhd.none_of_the_above.1 = NA.30.none_of_these.1,
    mhd.dont_know.1 = NA.30.dont_know.1,
    mhd.prefer_not_to_answer.1 = NA.30.prefer_not_to_answer.1
  ) %>%
  add_numeric(., exclude = exclude_cols) %>%
  mutate_if(is.numeric, ~na_if(., -88)) %>% # Recode missing values to NAs in numeric variables
  mutate_if(is.numeric, ~na_if(., -99)) %>%
  mutate_if(is.numeric, ~na_if(., -77)) %>%
  mutate_if(is.factor, ~na_if(., "Seen but not answered")) %>% # Recode missing values to NAs in factor variables
  mutate_if(is.factor, ~na_if(., "Don't know")) %>%
  mutate_if(is.factor, ~na_if(., "Prefer not to say")) %>%
  mutate_if(is.factor, ~na_if(., "Prefer not to answer")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Seen but not answered")) %>% # Drop empty factor levels
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Don't know")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to answer")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to say"))

@@

Agoraphobia and panic disorder group
```{r mhd agoraphobia_panic_disorder}
if(GLAD == TRUE | EDGI == TRUE | NBR == TRUE | RAMP == TRUE) {
mhd.raw.id <- mhd.raw.id %>%
  mutate(
    agoraphobia_panic_disorder_numeric =
      case_when(
        mhd.agoraphobia_numeric == "1" ~ 1,
        mhd.panic_disorder_numeric == "1" ~ 1,
        mhd.panic_attacks_numeric == "1" ~ 1,
        (mhd.agoraphobia_numeric == "0" &
        mhd.panic_disorder_numeric == "0" &
        mhd.panic_attacks_numeric == "0") ~ 0
      )
  )

@@

### Controls
These individuals have no self report diagnosis of any kind that we measure. 
```{r mhd control}
mhd.raw.id <- mhd.raw.id %>%
  mutate(
    control_numeric =
      if_else(
    depressive_disorders_numeric == "0" &
    anxiety_disorders_numeric == "0" &
    eating_disorders_numeric == "0" &
    obsessive_compulsive_disorders_numeric == "0" &
    psychotic_disorders_numeric == "0" &
    mhd.mania_hypomania_bipolar_or_manicdepression_numeric == "0" &
    mhd.posttraumatic_stress_disorder_ptsd_numeric == "0" &
    autism_spectrum_disorder_numeric == "0" &
    mhd.attention_deficit_hyperactivity_disorder_numeric == "0" &
    mhd.personality_disorder_numeric == "0" &
      mhd.prefer_not_to_answer_numeric == "0" & # added 31. August 2020
      mhd.prefer_not_to_answer.1_numeric == "0" & # added 31. August 2020
      mhd.dont_know_numeric == "0" & # added 31. August 2020
      mhd.dont_know.1_numeric == "0", # added 31. August 2020
    true = 1, 
    false = 0, 
    NA_real_)
  )

@@

```{r mhd check none of the above}
if(RAMP == TRUE) {
mhd.raw.id %>%
  select(
    control,
    mhd.none_of_the_above,
    mhd.none_of_the_above.1
  )

mhd.raw.id %>%
  group_by(control) %>%
  count(
    mhd.none_of_the_above.1,
    mhd.none_of_the_above
  )

#mhd.raw.id %>%
#  count(
#    control,
#    age_
#  )
}
```

Inspect the participants that do not qualify as double "None of these"
```{r RAMP none of these comparison}
if(RAMP == TRUE) {
mhd.raw.id %>%
  filter(
    (mhd.none_of_the_above.1 == "Not None of these" |
    mhd.none_of_the_above == "Not None of these") &
      control == "Control"
  ) %>%
    select(
      mhd.none_of_the_above_numeric,
      mhd.dont_know_numeric,
      mhd.prefer_not_to_answer_numeric,
      mhd.none_of_the_above.1_numeric,
      mhd.dont_know.1_numeric,
      mhd.prefer_not_to_answer.1_numeric
      )
}
```

```{r Check control columns}
if(GLAD == TRUE | NBR == TRUE | RAMP == TRUE) {
mhd.raw.id %>%
  select(
    control,
    depressive_disorders_numeric,
    anxiety_disorders_numeric,
    eating_disorders_numeric,
    obsessive_compulsive_disorders_numeric,
    psychotic_disorders_numeric,
    mhd.mania_hypomania_bipolar_or_manicdepression_numeric,
    mhd.posttraumatic_stress_disorder_ptsd_numeric,
    autism_spectrum_disorder_numeric,
    mhd.attention_deficit_hyperactivity_disorder_numeric,
    mhd.personality_disorder_numeric
  )
}
```

@@

Create new dataframe with the derived categories
```{r mhd categories}
if(GLAD == TRUE | NBR == TRUE | RAMP == TRUE) {
mhd.raw.categories <- mhd.raw.id %>%
  select(
    Sample,
    ID,
    anxiety_disorders_numeric,
    anxiety_disorders,
    mhd.specific_phobia_e.g._phobia_of_flying_numeric,
    mhd.specific_phobia_e.g._phobia_of_flying,
    mhd.social_anxiety_or_social_phobia_numeric,
    mhd.social_anxiety_or_social_phobia,
    agoraphobia_panic_disorder_numeric,
    agoraphobia_panic_disorder,
    depressive_disorders_numeric,
    depressive_disorders,
    eating_disorders_numeric,
    eating_disorders,
    obsessive_compulsive_disorders_numeric,
    obsessive_compulsive_disorders,
    psychotic_disorders_numeric,
    psychotic_disorders,
    mhd.mania_hypomania_bipolar_or_manicdepression_numeric,
    mhd.mania_hypomania_bipolar_or_manicdepression,
    mhd.posttraumatic_stress_disorder_ptsd_numeric,
    mhd.posttraumatic_stress_disorder_ptsd,
    autism_spectrum_disorder_numeric,
    autism_spectrum_disorder,
    mhd.attention_deficit_hyperactivity_disorder_numeric,
    mhd.attention_deficit_hyperactivity_disorder,
    mhd.personality_disorder_numeric,
    mhd.personality_disorder,
    personality_cluster_a_numeric,
    personality_cluster_a,
    personality_cluster_b_numeric,
    personality_cluster_b,
    personality_cluster_c_numeric,
    personality_cluster_c,
    depression_and_anxiety_numeric,
    depression_and_anxiety,
    restricting_vs_binge_eating_eating_disorders_numeric,
    restricting_vs_binge_eating_eating_disorders,
    bipolar_and_schizophrenia_numeric,
    bipolar_and_schizophrenia,
    control_numeric,
    control,
    disorders_total_count,
    comorbidity_total_count_numeric,
    comorbidity_total_count_factor_numeric,
    comorbidity_total_count_factor,
    mhd.none_of_the_above,
    mhd.none_of_the_above.1
    )
}
```

@@

```{r Rename dataset for ramp}
if(RAMP == TRUE){
#there is no sex variable in ramp
dat.no.sex <- dat.raw
}
```

@@

```{r Investigating RAMP NAs}
dat %>%
  group_by(control) %>%
  count(
    mhd.none_of_the_above.1,
    mhd.none_of_the_above,
    control,
    age_category
  )

dat %>%
  select(contains("none"))

dat %>%
  filter(
    (mhd.none_of_the_above.1 == "Not None of these" |
    mhd.none_of_the_above == "Not None of these") &
      control == "Control"
  )
